Low load operating protocol investigation of a 620MWe power boiler using an indirectly coupled process model

### 2021-09-27 PAPER_2/PAPER_2_REV1.tex
Update 27/09/2021

@@

Low load operation of utility boiler

\end{abstract}

@@

@@

\begin{center}
\begin{tabular}{|p{0.075\textwidth}p{0.3\textwidth}p{0.05\textwidth}p{0.075\textwidth}p{0.3\textwidth}p{0.05\textwidth}|} 
 \hline
\multicolumn{3}{|l}{\textbf{Nomenclature}} & &  &\\
\textit{Symbol} & \textit{Quantity} & \textit{Unit} & & &\\
$A$ &  Area & $m^2$ & & &\\
$A_p$ & Particle surface area & $m^2$& & & \\
$d_p$&Particle diameter & $m$ & & & \\
$E$& Fluid total energy&$J/kg$ & & & \\
$E_a$& Reaction activation energy&$J/kmol$ & & & \\
$P$& Pressure & $Pa$ & & & \\
$T_g$& Gas temperature& $K$ & & & \\
$T_p$&Particle temperature & $K$& & & \\
$u$& Velocity &$m/s^2$ & & &\\
\hline
\end{tabular}
\end{center}

\section{Introduction}
The use of coal fired power plants (CFPP) to provide electricity generation is intended to be phased out in order to mitigate the ingress of climate change. However, the switch to more sustainable generation sources pose great challenges for developing countries due to the long transition times and costs involved \cite{ugum2019}. Due to the abundance of coal resource present in South Africa CFPPs are the dominant power generation source, with approximately 80 $\%$ of the energy needs being met using CFPPs \cite{eskom}. The promising integration of renewable energy and the decommissioning of old CFPPs in South Africa, will push CFPP from a primarily base load operation to a mid-merit/flexible operating protocol. This will inherently mean CFPPs will need to operate at low-loads for continuous time periods.

@@

\section{Mathematical model}
In this section the modelling techniques used by the study are elaborated. A description of the CFD modelling configuration is discussed focusing on the fluid flow, turbulence and combustion modelling as well as the particle transport field resolution. Following this is a description of the heat transfer modelling techniques and the ends with a description of the process modelling configuration.
\subsection{Computational fluid dynamics modelling}
\subsubsection{Fluid flow, turbulence and combustion modelling}
The flue gas was modelled using a Eulerian framework. The species transport modelling approach was used to approximate the mixture of chemical species in the gas phase. This approach solves a species continuity equation for each constituent present in the mixture. To reduce the computational burden it was assumed that the various processes were in steady-state. The governing equations for the gas phase are written in their respective Reynolds averaged forms as follows;\\
Mass conservation:
\begin{equation}\label{eqn_RANS_mass}
\frac{\partial}{\partial x_{i}}(\rho \bar{u}_{i})=S_{m}.
\end{equation}
Momentum conservation:
\\
\begin{equation}\label{eqn_momentum}
\frac{\partial}{\partial x_{i}}(\rho_{eff} u_{i}u_{j})+\frac{\partial \overline{P}}{\partial x_{j}}=\frac{\partial}{\partial x_{i}}\left[\mu\left\{\frac{\partial u_{j}}{\partial x_{i}}+\frac{\partial u_{i}}{\partial x_{j}}-\frac{2}{3}\delta_{ij}\frac{\partial u_{i}}{\partial x_{i}}\right\}\right]+\frac{\partial}{\partial x_{i}}(-\rho\overline{u_{i}^{'}u_{j}^{'}})+S_m
\end{equation}\\
\\
Energy conservation:
\begin{equation}\label{eqn_energy}
\frac{\partial }{\partial x_{i}} (u_{i}[\rho E+P])=\frac{\partial }{\partial x_{j}}\left[\lambda\frac{\partial T_{g}}{\partial x_{j}}\right] +S_{h}
\end{equation}
\\
Species transport:
\begin{equation}\label{eqn_species}
\begin{split}
&\frac{\partial}{\partial x_{i}}(\rho u_{j}Y_{k})=-\frac{\partial}{\partial x_{j}}(\vec{J_{k}})+ \sum_r R_{j,r} + S_{k}\\
&k = 1,2,3...N
\end{split}
\end{equation}

To correctly account for the particle inertial effects on the gas phase convection an effective density is defined as follows;
\begin{equation} \label{eqn_eff_rho}
	\rho_{eff} = \frac{\rho \rho_p \left( \phi_{mp} + 1 \right)}{\rho \phi_{mp} + \rho_p}
\end{equation}

@@

A devolatilization temperature of 553 [$K$] \citep{Ranade2015} along with the kinetic parameters (equation \ref{eqn_vol_arrhenuis}) of Sheng et al \cite{Sheng2004} were utilized. The char oxidation process is modelled using the diffusion-kinetics limited model developed by Baum and Street \cite{Baum1971}, which is given in equation (\ref{eqn_char_rate}). The product species of the char oxidation reaction was set to $CO$ as shown in equation (\ref{eqn_CO_reaction}). 
\begin{gather}
\frac{dm_{char}}{dt} = -A_p P_{O_{2}} \frac{R_{diff}R_c}{R_{diff} + R_c}  \label{eqn_char_rate}\\
C_{(s)}+0.5O_{2(g)}\to CO_{(g)} \label{eqn_CO_reaction}
\end{gather}
The diffusion and kinetic rates of equation (\ref{eqn_char_rate}) are defined in equations (\ref{eqn_char_diff_rate})  and (\ref{eqn_char_kin_rate}) with the kinetic parameters again taken from the works of Sheng et al \citep{Sheng2004}.
\begin{gather}
R_{diff} = \frac{5X10^{-12}}{d_p} \left(\frac{T_g+T_p}{2}\right)^0.75 \label{eqn_char_diff_rate}\\
\begin{split}
&R_{c} = A_{c}exp\left(\frac{E_{a,c}}{RT_p}\right)\\
&A_{c} = 0.0053 [kg/m^2sPa]\,\,\,\,\,\,\,\,\,\,E_{a,c} = 8.37\times10^7 [J/kmol]
\end{split}
 \label{eqn_char_kin_rate}
\end{gather}

@@

The energy transport of the pseudo particle, is transported by defining the particle enthalpy using the following equation:
\begin{equation}\label{eqn_phi_hp}
\frac{\partial}{\partial x_{i}}(\rho u_{i} \phi_{hp})=\left(f_{heat}\frac{dM_{c}}{dt}h_{rxn} + \dot{Q}_{rad} + \dot{Q}_{conv} - \frac{dM_{evap}}{dt}h_{fg}\right)\frac{1}{V}
\end{equation}

The equation accounts for all the processes associated with energy transport to the particle, namely convection $\left(\dot{Q}_{conv}\right)$, radiation $\left(\dot{Q}_{rad}\right)$, latent heat $\left(\frac{dM_{evap}}{dt}h_{fg}\right)$ and near surface char oxidation $\left(f_{heat}\frac{dM_{c}}{dt}h_{rxn}\right)$. This gives the model the ability to track the particle temperature in the domain, moving the model away from the thermal equilibrium approach incorporated by previous studies using an EE approach (\cite{Benim2005}, \cite{Vicente2003} and \cite{Cai2015}). The particle temperature is important in describing the sequential steps found in modelling combustion processes, especially at low boiler loads where mixing and ignition become problematic.

\subsection{Heat transfer modelling modelling}
The radiation heat transfer is the dominant form of heat transfer found in industrial furnaces \citep{Basu2000} and is solved by applying the gray-participating-gas and particle medium configuration of the radiation transport equation (RTE) \cite{Modest2013} shown in equation (\ref{eqn_RTE}).
\begin{equation}\label{eqn_RTE}
\frac{d I(\vec{r},\hat{s})}{ds} = \alpha_g \frac{\sigma_SB T_{g}^4}{\pi}-(\alpha_g+\alpha_p+\sigma_p)I(\vec{r},\hat{s}) + \frac{\sigma_p}{4\pi}\int_{4\pi}I(\vec{r},\hat{s})\Phi d \Omega
\end{equation}
In the present work the RTE is solved sing the P1 model. Ranade and Gupta \cite{Ranade2015} illustrated minimal differences between the two common radiation models (namely the P1 and discrete ordinates (DO)) for the resultant wall heat transfer rate values when modelling a 210 MWe CFPP boiler. The P1 radiation model can include the effects of particle absorption ($\alpha_p$) and scattering ($\sigma_p$) as well as gas mixture absorption ($\alpha_g$). The P1 model transport variable is the incident radiation (G - [$W/m^2$]), and can be written for a particle laden domain as:
\begin{equation}
\begin{split}
&\frac{\partial}{\partial x_{i}}\left(\Gamma\frac{\partial G}{\partial x_{i}}\right)=\left(\alpha_g+\alpha_p\right)G-4\left(\alpha_g \sigma_{SB} T_{g}^4-\pi E_p \right)\\
&\Gamma = \frac{1}{\alpha_g+\alpha_p+\sigma_p}
\end{split}
\end{equation}

The flue gas absorptivity was calculated using the domain based weighted sum of gray gas model (WSGGM) using the coefficients determined by Smith et al \cite{Smith1982}. The WSGGM accounts for the radiation emitted by tri-atomic gases, namely $CO_2$, $H_2O$ and $SO_2$ present in the flue gas stream. The Eulerian description of the terms $\alpha_p$, $\sigma_p$ and $E_b$ are determined using the effective number of particles ($N_p$) present in a cell. There formulations are given in equations (\ref{eqn_part_abs}) through (\ref{eqn_part_emisp}).
\begin{gather}
\alpha_p = \frac{\epsilon_p A_{pn}N_p}{V} \label{eqn_part_abs}\\
\sigma_p = \frac{(1-\epsilon_p)(1-f_p) A_{pn}N_p}{V} \label{eqn_part_scat} \\
E_p = \frac{\epsilon_p \sigma_{SB} T_p^4 A_{pn}N_p}{V}\label{eqn_part_emisp}
\end{gather}

It is important to note that variable properties for ($\epsilon_p$) and ($f_p$) are used, that are based on the correlations of Lockwood et al \cite{Lockwood1986} and Yin \citep{Yin2015} respectively.

@@

\begin{table}[h!]
\centering
\caption{Case 1 to 6 model inputs on a per burner basis.}
\label{tbl_case_inputs}
\vspace{5mm}
\label{fuel}
{\tabulinesep=1.2mm
\begin{tabularx}{\textwidth}{p{0.45\textwidth} p{0.25\textwidth} l}
\hline
Active burners & \textbf{Cases 1 - 3} & \textbf{Cases 4 - 6}\\
\hline
\textbf{Fuel flow rate} [$kg/s$]&3.14  &3.14\\
\textbf{PA flow rate} [$kg/s$]&4.95  &4.95\\
\textbf{SA flow rate} [$kg/s$]&14.85  &14.85\\
\hline
Non firing burners &  & \\
\hline
\textbf{SA flow rate} [$kg/s$]&5.0  &2.5\\
\hline
Input air temperatures& &\\
\hline
\textbf{PA} [$K$]&373  &373\\
\textbf{SA} [$K$]&520  &510\\
\hline
\end{tabularx}}
\end{table}

Table \ref{tbl_case_inputs} shows a lower mass flow rate of SA air for case 4 to 6, this was done to compare the expected increase in boiler utilization efficiency, since less dry gas losses are expected, and the effects the drop in the flue-gas mass flow rate has on the convective pass components modelled in the process model.

@@

The spatial discretization for all fields (except pressure) was set to first-order upwind for the first 1000 iterations to ensure a stable solution, after which the discretization order was increased to the above mentioned criteria. For all cases the maximum mass imbalance was 0.024 [$kg/s$] for a total gas flow rate of 190 [$kg/s$] and a heat imbalance of 1770 [$kW$] for a total heat input of 283 [$MW$]. The remaining fields were solved till convergence.

@@

\begin{figure}[h!]
	\subfigure[]{\includegraphics[width=0.32\textwidth]{100_VALID_BAR}}
	\subfigure[]{\includegraphics[width=0.32\textwidth]{80_VALID_BAR}}
	\subfigure[]{\includegraphics[width=0.32\textwidth]{60_VALID_BAR}}
\caption{Comparison of experimentally calculated and model heat loads for the furnace, platen SH and pendant SH at (a) 100\% MCR, (b) 80\% MCR and (c) 60\% MCR}
\label{fig_heat_valid}
\end{figure}
\clearpage
In Figure \ref{fig_heat_valid} it is shown that the overall heat loads are in good agreement with the calculated results. For the simulated validation loads the proposed model results are within the associated error band, the general trend is an under prediction on the furnace heat loads and an over prediction on the platen super-heater. The pendant super heater illustrate the best comparable results for all load cases.

The CFD model was further validated by comparing the $CO_{ppm}$ and $X_{O_{2}}$ measurements against the CFD results. The probe measurements were taken at a furnace height of 37.5 [$m$] near the center of the boiler during a full load (100\% MCR) operating conditions. The probe is inserted from the side walls to a depth of 4.5 [$m$], measurements were taken every 0.5 [$m$].\\
\begin{figure}[h!]
\centering
\subfigure[]{\includegraphics[width = 0.45\textwidth, height =4cm ]{COPPM_VALID}}
\hspace{5mm}
\subfigure[]{\includegraphics[width = 0.45\textwidth, height =4cm ]{XO2_VALID}}
\caption{Experimentally calculated $CO_{ppm}$ (a) and $X_{O_{2}}$ (b) concentration predictions}
\label{fig_probe_valid}
\end{figure}

Figure \ref{fig_probe_valid} shows the averaged measurement values to that of the CFD predictions. It can be seen that the CFD model can sufficiently resolve the $CO_{ppm}$ and $X_{O_{2}}$ concentrations at the given probe location.
\newpage
\subsection{Simulation results for various burner firing arrangements at 32\% MCR }

Figures \ref{fig_cfd_temp} and \ref{fig_cfd_velo} illustrate the temperature and velocity profiles for the various firing arrangements. With only the middle burner rows firing (Cases 2 and 5) there a is substantial cold region located in the lower half the furnace, resulting in the lowest heat uptake of the furnace for all the cases. This is further exaggerated when considering the Table \ref{tbl_process_parameters} with the mid-firing arrangement producing the lowest steam flow rates resulting in the lowest boiler efficiencies. The bottom firing arrangement (Cases 1 and 4) results in fireball located in the bottom half on the burner, while the mixed firing arrangement (Cases 3 and 6) results is more even distribution of the fireball in the furnace domain. This leads to the highest steam generation rate and boiler efficiency as shown in Table \ref{tbl_process_parameters}.

Figure \ref{fig_cfd_heat_flux} illustrates the heat flux profiles for the simulated cases. Both cases 1 and 4 highlight high heat flux zones near the burner inlets which, in the presence of high temperatures and incomplete combustion near these regions, can lead to high-temperature corrosion \cite{Du2017}. Cases 2 and 5 show that most of the unit heat fluxes are absorbed in the upper half of the boiler, contrary to cases 1 and 4, where the lower half has a higher heat flux absorption, and cases 3 and 6 showing an even distribution throughout the furnace.

>>>>

\usepackage{framed}
\usepackage{multicol} % Multiple columns environment

\usepackage{nomencl} % Nomenclature package
\makenomenclature
\setlength{\nomitemsep}{0pt}

\newcommand{\nomunit}[1]{%
  \renewcommand{\nomentryend}{\hspace*{\fill}#1}%
}

\renewcommand\nomgroup[1]{% 
  \item[\bfseries
  \ifstrequal{#1}{A}{Symbols}{%
  \ifstrequal{#1}{B}{Greek symbols}{%
  \ifstrequal{#1}{C}{Abbreviations}{}}}%
]}

@@

Low load operation of utility boiler
ghfbksbv\\
sfjvnsfjlvl\\
kfnmvknsfklvns\\
lk;sfvnsfnvsd\\
sk,vnsfk;nvk;sdv\\
ksdnv;sdv\\
sknvk
\end{abstract}

@@

\printnomenclature

\nomenclature[A]{$A$}{Area \nomunit{$m^2$}}
\nomenclature[A]{$A_p$}{Particle surface area \nomunit{$m^2$}}
\nomenclature[A]{$A_{pn}$}{Projected particle surface area \nomunit{$m^2$}}
\nomenclature[A]{$A_{c}$}{Carbon pre-exponential factor \nomunit{$s^{-1}$}}
\nomenclature[A]{$A_{vol}$}{Volatile pre-exponential factor \nomunit{$s^{-1}$}}
\nomenclature[A]{$d_p$}{Particle diameter \nomunit{$m$}}
\nomenclature[A]{$E$}{Fluid total energy \nomunit{$J/kg$}}
\nomenclature[A]{$E_{a,c}$}{Carbon activation energy \nomunit{$J/kmol$}}
\nomenclature[A]{$E_{a,vol}$}{Volatile activation energy \nomunit{$J/kmol$}}
\nomenclature[A]{$E_p$}{Particle emission power \nomunit{$W/m^3$}}
\nomenclature[A]{$f_{heat}$}{Near surface char oxidation fraction \nomunit{$-$}}
\nomenclature[A]{$f_{p}$}{Forward scattering factor \nomunit{$-$}}
\nomenclature[A]{$G$}{Incident radiation \nomunit{$W/m^2$}}
\nomenclature[A]{$g$}{Gravitational constant \nomunit{$m/s^2$}}
\nomenclature[A]{$h_{fg}$}{Enthalpy of evaporation \nomunit{$J/kg$}}
\nomenclature[A]{$h_M$}{Mixture enthalpy \nomunit{$J/kg$}}
\nomenclature[A]{$h_{rxn}$}{Enthalpy of reaction \nomunit{$J/kg$}}
\nomenclature[A]{$J_k$ }{Species diffusion flux \nomunit{$m^2/s$}}
\nomenclature[A]{$m_{0,vol}$}{Original mass of volatiles \nomunit{$kg$}}
\nomenclature[A]{$m_{char}$}{Mass of carbon \nomunit{$kg$}}
\nomenclature[A]{$m_{evap}$}{Mass of moisture \nomunit{$kg$}}
\nomenclature[A]{$m_{vol}$}{Mass of volatiles \nomunit{$kg$}}
\nomenclature[A]{$N_{p}$}{Number of particles \nomunit{$-$}}
\nomenclature[A]{$P$}{Perimeter \nomunit{$m$}}
\nomenclature[A]{$p$}{Pressure \nomunit{$Pa$}}
\nomenclature[A]{$p_{O_{2}}$}{Partial pressure of $O_2$ \nomunit{$Pa$}}
\nomenclature[A]{$\dot{Q}_{conv}$}{Convective heat transfer rate \nomunit{$W$}}
\nomenclature[A]{$\dot{Q}_{rad}$}{Radiative heat transfer rate \nomunit{$W$}}
\nomenclature[A]{$\dot{Q}_{w}$}{Wall heat transfer rate \nomunit{$W$}}
\nomenclature[A]{$R$}{Universal gas constant \nomunit{$J/molK$}}
\nomenclature[A]{$R_{c}$}{Carbon reaction rate \nomunit{$s^{-1}$}}
\nomenclature[A]{$R_{diff}$}{Diffusion reaction rate \nomunit{$s^{-1}$}}
\nomenclature[A]{$R_{vol}$}{Volatile reaction rate \nomunit{$s^{-1}$}}
\nomenclature[A]{$T_g$}{Gas temperature \nomunit{$K$}}
\nomenclature[A]{$T_p$}{Particle temperature \nomunit{$K$}}
\nomenclature[A]{$t$}{Time \nomunit{$s$}}
\nomenclature[A]{$u$}{Velocity \nomunit{$m/s$}}
\nomenclature[A]{$V$}{Volume \nomunit{$m^3$}}
\nomenclature[A]{$Y_k$}{Species mass fraction \nomunit{$kg/kg$}}
\nomenclature[A]{$x$}{Quality \nomunit{$-$}}
\nomenclature[B]{$\alpha_g$}{Gas absorption coefficient \nomunit{$m^{-1}$}}
\nomenclature[B]{$\alpha_H$}{Mixture void fraction \nomunit{$-$}}
\nomenclature[B]{$\alpha_p$}{Particle absorption coefficient \nomunit{$m^{-1}$}}
\nomenclature[B]{$\epsilon$}{Emissivity \nomunit{$-$}}
\nomenclature[B]{$\rho$}{Gas density \nomunit{$kg/m^3$}}
\nomenclature[B]{$\rho_{eff}$}{Effective density \nomunit{$kg/m^3$}}
\nomenclature[B]{$\rho_g$}{Gas density \nomunit{$kg/m^3$}}
\nomenclature[B]{$\rho_l$}{Liquid density \nomunit{$kg/m^3$}}
\nomenclature[B]{$\rho_M$}{Mixture density \nomunit{$kg/m^3$}}
\nomenclature[B]{$\rho_p$}{Particle density \nomunit{$kg/m^3$}}
\nomenclature[B]{$\sigma_{SB}$}{Stefan-Boltzmann constant \nomunit{$W/m^2K^4$}}
\nomenclature[B]{$\sigma_{p}$}{Particle scattering coefficient \nomunit{$m^{-1}$}}
\nomenclature[B]{$\phi$}{Scalar variable \nomunit{$-$}}
\nomenclature[C]{$MCR$}{Maximum continuous rating}
\nomenclature[C]{$CFPP$}{Coal fired power plant}
\nomenclature[C]{$CFD$}{Computational fluid dynamics}
\nomenclature[C]{$SH$}{Superheater}
\nomenclature[C]{$EE$}{Eulerian-Eulerian}
\nomenclature[C]{$WSGGM$}{Weighted sum of gray gas model}
\nomenclature[C]{$RTE$}{Radiative transport equation}
\nomenclature[C]{$HHV$}{Higher heating value}
\nomenclature[C]{$DAF$}{Dry ash free}
\nomenclature[C]{$AR$}{As-received}
\nomenclature[C]{$PA$}{Primary air}
\nomenclature[C]{$SA-AH$}{Secondary air - air heater}
\nomenclature[C]{$SA$}{Secondary air}
\nomenclature[C]{$RH$}{Reheater}
\nomenclature[C]{$ATT$}{Attemperator}
\nomenclature[C]{$EC$}{Economiser}
\nomenclature{\mbox{}}{}

@@

\section{Introduction}
The use of coal fired power plants (CFPP) to provide electricity generation is intended to be phased out in order to mitigate the ingress of climate change. However, the switch to more sustainable generation sources pose great challenges for developing countries due to the long transition times and costs involved \cite{ugum2019}. Due to the abundance of coal resource present in South Africa CFPPs are the dominant power generation source, with approximately 80 $\%$ of the energy needs being met using CFPPs \cite{eskom}. The promising integration of renewable energy and the decommissioning of old CFPPs in South Africa, will push CFPP from a primarily base load operation to a mid-merit/flexible operating protocol. This will inherently mean CFPPs will need to operate at low-loads for continuous time periods.

@@

\section{Mathematical model}
In this section the modelling techniques used by the study are elaborated. A description of the CFD modelling configuration is discussed focusing on the fluid flow, turbulence and combustion modelling as well as the particle transport field resolution. Following this is a description of the heat transfer modelling techniques and the ends with a description of the process modelling configuration.
\subsection{Computational fluid dynamics modelling}
\subsubsection{Fluid flow, turbulence and combustion modelling}
The flue gas was modelled using a Eulerian framework. The species transport modelling approach was used to approximate the mixture of chemical species in the gas phase. This approach solves a species continuity equation for each constituent present in the mixture. To reduce the computational burden it was assumed that the various processes were in steady-state. The governing equations for the gas phase are written in their respective Reynolds averaged forms as follows;\\
Mass conservation:
\begin{equation}\label{eqn_RANS_mass}
\frac{\partial}{\partial x_{i}}(\rho \bar{u}_{i})=S_{m}.
\end{equation}
Momentum conservation:
\\
\begin{equation}\label{eqn_momentum}
\frac{\partial}{\partial x_{i}}(\rho_{eff} u_{i}u_{j})+\frac{\partial \overline{p}}{\partial x_{j}}=\frac{\partial}{\partial x_{i}}\left[\mu\left\{\frac{\partial u_{j}}{\partial x_{i}}+\frac{\partial u_{i}}{\partial x_{j}}-\frac{2}{3}\delta_{ij}\frac{\partial u_{i}}{\partial x_{i}}\right\}\right]+\frac{\partial}{\partial x_{i}}(-\rho\overline{u_{i}^{'}u_{j}^{'}})+S_m
\end{equation}\\
Energy conservation:
\begin{equation}\label{eqn_energy}
\frac{\partial }{\partial x_{i}} (u_{i}[\rho E+p])=\frac{\partial }{\partial x_{j}}\left[\lambda\frac{\partial T_{g}}{\partial x_{j}}\right] +S_{h}
\end{equation}
\\
Species transport:
\begin{equation}\label{eqn_species}
\begin{split}
&\frac{\partial}{\partial x_{i}}(\rho u_{j}Y_{k})=-\frac{\partial}{\partial x_{j}}(\vec{J_{k}})+ \sum_r R_{j,r} + S_{k}\\
&k = 1,2,3...N
\end{split}
\end{equation}

To correctly account for the particle inertial effects on the gas phase convection, the model makes use of an effective density is defined present as follows;
\begin{equation} \label{eqn_eff_rho}
	\rho_{eff} = \frac{\rho \rho_p \left( \phi_{mp} + 1 \right)}{\rho \phi_{mp} + \rho_p}
\end{equation}

@@

A devolatilization temperature of 553 [$K$] \citep{Ranade2015} along with the kinetic parameters (equation \ref{eqn_vol_arrhenuis}) of Sheng et al \cite{Sheng2004} were utilized. The char oxidation process is modelled using the diffusion-kinetics limited model developed by Baum and Street \cite{Baum1971}, which is given in equation (\ref{eqn_char_rate}). The product species of the char oxidation reaction was set to $CO$ as shown in equation (\ref{eqn_CO_reaction}). 
\begin{gather}
\frac{dm_{char}}{dt} = -A_p p_{O_{2}} \frac{R_{diff}R_c}{R_{diff} + R_c}  \label{eqn_char_rate}\\
C_{(s)}+0.5O_{2(g)}\to CO_{(g)} \label{eqn_CO_reaction}
\end{gather}
The diffusion and kinetic rates of equation (\ref{eqn_char_rate}) are defined in equations (\ref{eqn_char_diff_rate})  and (\ref{eqn_char_kin_rate}) with the kinetic parameters again taken from the works of Sheng et al \citep{Sheng2004}.
\begin{gather}
R_{diff} = \frac{5\times10^{-12}}{d_p} \left(\frac{T_g+T_p}{2}\right)^{0.75} \label{eqn_char_diff_rate}\\
\begin{split}
&R_{c} = A_{c}exp\left(\frac{E_{a,c}}{RT_p}\right)\\
&A_{c} = 0.0053 [kg/m^2sPa]\,\,\,\,\,\,\,\,\,\,E_{a,c} = 8.37\times10^7 [J/kmol]
\end{split}
 \label{eqn_char_kin_rate}
\end{gather}

@@

The energy transport of the pseudo particle, is transported by defining the particle enthalpy using the following equation:
\begin{equation}\label{eqn_phi_hp}
\frac{\partial}{\partial x_{i}}(\rho u_{i} \phi_{hp})=\left(f_{heat}\frac{dm_{c}}{dt}h_{rxn} + \dot{Q}_{rad} + \dot{Q}_{conv} - \frac{dm_{evap}}{dt}h_{fg}\right)\frac{1}{V}
\end{equation}

The equation accounts for all the processes associated with energy transport to the particle, namely convection $\left(\dot{Q}_{conv}\right)$, radiation $\left(\dot{Q}_{rad}\right)$, latent heat $\left(\frac{dm_{evap}}{dt}h_{fg}\right)$ and near surface char oxidation $\left(f_{heat}\frac{dm_{c}}{dt}h_{rxn}\right)$. This gives the model the ability to track the particle temperature in the domain, moving the model away from the thermal equilibrium approach incorporated by previous studies using an EE approach (\cite{Benim2005}, \cite{Vicente2003} and \cite{Cai2015}). The particle temperature is important in describing the sequential steps found in modelling combustion processes, especially at low boiler loads where mixing and ignition become problematic.

\subsection{Heat transfer modelling modelling}
The radiation heat transfer is the dominant form of heat transfer found in industrial furnaces \citep{Basu2000} and is solved by applying the gray-participating-gas and particle medium configuration of the radiation transport equation (RTE) \cite{Modest2013} shown in equation (\ref{eqn_RTE}).
\begin{equation}\label{eqn_RTE}
\frac{d I(\vec{r},\hat{s})}{ds} = \alpha_g \frac{\sigma_{SB} T_{g}^4}{\pi}-(\alpha_g+\alpha_p+\sigma_p)I(\vec{r},\hat{s}) + \frac{\sigma_p}{4\pi}\int_{4\pi}I(\vec{r},\hat{s})\Phi d \Omega
\end{equation}
In the present work the RTE is solved sing the P1 model. Ranade and Gupta \cite{Ranade2015} illustrated minimal differences between the two common radiation models (namely the P1 and discrete ordinates (DO)) for the resultant wall heat transfer rate values when modelling a 210 MWe CFPP boiler. The P1 radiation model can include the effects of particle absorption ($\alpha_p$) and scattering ($\sigma_p$) as well as gas mixture absorption ($\alpha_g$). The P1 model transport variable is the incident radiation (G - [$W/m^2$]), and can be written for a particle laden domain as:
\begin{equation}
\begin{split}
&\frac{\partial}{\partial x_{i}}\left(\Gamma\frac{\partial G}{\partial x_{i}}\right)=\left(\alpha_g+\alpha_p\right)G-4\left(\alpha_g \sigma_{SB} T_{g}^4-\pi E_p \right)\\
&\Gamma = \frac{1}{\alpha_g+\alpha_p+\sigma_p}
\end{split}
\end{equation}

The flue gas absorptivity was calculated using the domain based weighted sum of gray gas model (WSGGM) using the coefficients determined by Smith et al \cite{Smith1982}. The WSGGM accounts for the radiation emitted by tri-atomic gases, namely $CO_2$, $H_2O$ and $SO_2$ present in the flue gas stream. The Eulerian description of the terms $\alpha_p$, $\sigma_p$ and $E_b$ are determined using the effective number of particles ($N_p = \rho \phi_{mp0} V / \left( \rho_p \pi \bar{d_p}^3 /6 \right)$) present in a cell. There formulations are given in equations (\ref{eqn_part_abs}) through (\ref{eqn_part_emisp}).
\begin{gather}
\alpha_p = \frac{\epsilon_p A_{pn}N_p}{V} \label{eqn_part_abs}\\
\sigma_p = \frac{(1-\epsilon_p)(1-f_p) A_{pn}N_p}{V} \label{eqn_part_scat} \\
E_p = \frac{\epsilon_p \sigma_{SB} T_p^4 A_{pn}N_p}{V}\label{eqn_part_emisp}
\end{gather}

It is important to note that variable properties for ($\epsilon_p$) and ($f_p$) were used, that are based on the correlations of Lockwood et al \cite{Lockwood1986} and Yin \citep{Yin2015} respectively.

@@

\begin{table}[h!]
\centering
\caption{Case 1 to 6 model inputs on a per burner basis.}
\label{tbl_case_inputs}
\vspace{2mm}
{\tabulinesep=1.2mm
\begin{tabularx}{\textwidth}{p{0.45\textwidth} p{0.25\textwidth} l}
\hline
Active burners & \textbf{Cases 1 - 3} & \textbf{Cases 4 - 6}\\
\hline
\textbf{Fuel flow-rate} [$kg/s$]&3.14  &3.14\\
\textbf{PA flow-rate} [$kg/s$]&4.95  &4.95\\
\textbf{SA flow-rate} [$kg/s$]&14.85  &14.85\\
\hline
Non firing burners &  & \\
\hline
\textbf{SA flow-rate} [$kg/s$]&5.0  &2.5\\
\hline
Input air temperatures& &\\
\hline
\textbf{PA} [$K$]&373  &373\\
\textbf{SA} [$K$]&520  &510\\
\hline
\end{tabularx}}
\end{table}

Table \ref{tbl_case_inputs} shows a lower mass flow rate of SA air for case 4 to 6, this was done to compare the expected increase in boiler utilization efficiency, since less dry gas losses are expected, and the effects the drop in the flue-gas mass flow-rate has on the convective pass components modelled in the process model.

@@

The spatial discretization for all fields (except pressure) was set to first-order upwind for the first 1000 iterations to ensure a stable solution, after which the discretization order was increased to the above mentioned criteria. For all cases the maximum mass imbalance was 0.024 [$kg/s$] for a total gas flow-rate of 190 [$kg/s$] and a heat imbalance of 1770 [$kW$] for a total heat input of 283 [$MW$]. The remaining fields were solved till convergence.

@@

\begin{figure}[h!]
	\subfigure[100 \% MCR]{\includegraphics[width=0.32\textwidth]{100_VALID_BAR}}
	\subfigure[80 \% MCR]{\includegraphics[width=0.32\textwidth]{80_VALID_BAR}}
	\subfigure[60 \% MCR]{\includegraphics[width=0.32\textwidth]{60_VALID_BAR}}
\caption{Comparison of experimentally calculated and model heat loads for the furnace, platen SH and final SH}
\label{fig_heat_valid}
\end{figure}
\clearpage
In Figure \ref{fig_heat_valid} it is shown that the overall heat loads are in good agreement with the calculated results. For the simulated validation loads the proposed model results are within the associated error band, the general trend is an under prediction on the furnace heat loads and an over prediction on the platen super-heater. The pendant super heater illustrate the best comparable results for all load cases.

The CFD model was further validated by comparing the $CO_{ppm}$ and $X_{O_{2}}$ measurements against the CFD results. The probe measurements were taken at a furnace height of 37.5 [$m$] near the center of the boiler during a full load (100\% MCR) operating conditions. The probe is inserted from the side walls to a depth of 4.5 [$m$], measurements were taken every 0.5 [$m$].\\
\begin{figure}[h!]
\centering
\subfigure[]{\includegraphics[width = 0.45\textwidth, height =4cm ]{COPPM_VALID}}
\hspace{5mm}
\subfigure[]{\includegraphics[width = 0.45\textwidth, height =4cm ]{XO2_VALID}}
\caption{Experimentally calculated $CO_{ppm}$ (a) and $X_{O_{2}}$ (b) concentration predictions}
\label{fig_probe_valid}
\end{figure}
Figure \ref{fig_probe_valid} shows the averaged measurement values compared to the CFD predictions. It can be seen that the CFD model can sufficiently resolve the $CO_{ppm}$ and $X_{O_{2}}$ concentrations at the given probe location.
\newpage
\subsection{Simulation results for various burner firing arrangements at 32\% MCR }
The temperature and velocity profiles for the various firing arrangements are shown in Figures \ref{fig_cfd_temp} and \ref{fig_cfd_velo} respectively. With only the middle burner rows firing (Cases 2 and 5) a substantial cold region is formed in the lower half the furnace, resulting in the lowest heat uptake. This is further exaggerated when considering Table \ref{tbl_process_parameters} where it is illustrated that the mid-firing arrangements produce the lowest steam flow rates resulting in the lowest boiler efficiencies. The bottom firing arrangement (Cases 1 and 4) results in a high temperature zone located in the bottom half on the burner, while the mixed firing arrangement (Cases 3 and 6) results is an even distribution of high temperature gases across the furnace domain (Figure \ref{fig_cfd_temp}). This leads to the highest steam generation rate and boiler efficiency as shown in Table \ref{tbl_process_parameters}.

\begin{figure}[h!]
\centering
\includegraphics[scale = 0.45]{TEMP_KEY} [$K$]\\
\subfigure[Case 1]{\includegraphics[width=0.32\textwidth]{BOT_TEMP}}
\subfigure[Case 2]{\includegraphics[width=0.32\textwidth]{MID_TEMP}}
\subfigure[Case 3]{\includegraphics[width=0.32\textwidth]{FBRM_TEMP}}\\
\subfigure[Case 4]{\includegraphics[width=0.32\textwidth]{BOT05_TEMP}}
\subfigure[Case 5]{\includegraphics[width=0.32\textwidth]{MID05_TEMP}}
\subfigure[Case 6]{\includegraphics[width=0.32\textwidth]{FBRM05_TEMP}}
\caption{Temperature fields for cases 1 through 6 [(a)-(f)]}
\label{fig_cfd_temp}
\end{figure}
